|

### 2023-09-25 chapter_2/src/roberts_empirical_chapter_1.qmd
integrated a bunch of them, not sure about the others. Will keep it open for now

# Does color convey political information?
<!--
    - Description: QMD script for empirical chapter 1
    - Updated: 2023-07-16
    - Notes:
        * Code:
        * Writing:
-->

@@

In [Study 1](#study-1), I conduct an online experiment that examines whether the public notice color and whether their attitudes are shaped by "Republican red" and "Democratic Blue" on the yard signs of a fictional candidate. Based on informal conversations with campaign marketing consultants, it appears that campaigns are aware that the colors red and blue betray the partisanship of a candidate to voters. Accordingly, in [Study 2](#study-2), I examine whether campaigns react to the partisan composition of an electorate, as a response to these individual-level patterns. Specifically, I do this by examining how electoral context shapes variation in the use of these partisan colors. While these two studies highlight endogenous processes, to be clear, I am not arguing that either of these patterns only one way. In fact, I believe that there is a necessary feedback loop between the parties and their bases of support. Voters support candidates who signal strong allegiance to the party to increase their chances for electoral success [@utych_voter-centric_2020], and the parties are motivated to distinguish themselves [@lee_insecure_2016;@clifford_how_2020] as a result of their high levels of ideological and affective polarization [@dietrich_using_2021;@enders_issues_2021]. I leave teasing out directionality for subsequent analyses.

In the [first study](#study-1), I find that individuals do shift their perceptions of political candidates in response to the use of the colors red and blue -- they consistently and strongly associate Republicans with the color red and Democrats with the color blue, and their willingness to vote for a candidate can be dramatically different depending on whether the color used aligns with their partisan identification. The [second study](#study-2) suggests that candidates in districts where Democrats have a higher vote share use more blue on their yard signs (while also using less red).

@@

```{r}
#| label: study-1-setup

The purpose of [Study 1](#sec-study-1) is to examine whether the color choices on an almost ubiquitous form of campaign branding -- yard signs -- influence people's perceptions of political candidates. Evidence suggests that yard signs *do* matter in shaping political attitudes [@makse_politics_2019], vote intentions [@makse_politics_2019], and even electoral outcomes [@green_effects_2016]. As they are simple, cheap, intrusive, and common forms of campaign branding, they provide a conservative test of the effects of color in campaign branding.

Using yard signs in this study, I test the first four hypotheses I derive from the snap-judgment model. First, I test the claim that individuals do indeed notice the color of electoral yard signs. Next, I test the claim that these colors that individuals detect influence their evaluations of the yard sign and the candidate represented on it. I then test the claim that the effects on perception are moderated by how consistent the color is with the more complex information displayed on the yard sign. Finally, [Study 1](#study-1) closes with an examination of whether positive information that contains consistency is processed more quickly than negative or inconsistent information. That is, do partisans quickly detect and encode information from a clearly co-partisan political candidate?

I recruit participants from Prolific.[^color_influence_pre-reg_anon-3] After providing informed consent to participate in the study, Prolific redirects subjects to Pavlovia,[^color_influence_pre-reg_anon-4] where individuals are provided with a demographics and political attitudes questionnaire. As I am concerned about priming effects introduced by these questions, while I am simultaneously concerned about bias introduced by post-treatment control [@montgomery_how_2018], I randomly select half of the participants to receive the questionnaire post-treatment and the other half receive it pre-treatment. This questionnaire includes common questions about the participant's ascriptive characteristics, and about the participant's political ideology, partisan identification, interest in politics, patriotism, and political knowledge.

I include those questions due to expectations that they may act as confounds for my hypotheses. I include some questions collecting information on participants' ascriptive and descriptive characteristics such as age, education, gender identity, and racial identity, as a number of these correlates with partisan identification [see @campbell_american_1969;@mason_uncivil_2018]. Though political knowledge does predict the strength of an individual's partisan identification [@lodge_rationalizing_2013], attention paid to politics is likely a more important confounding variable for my application. The purpose of color in politics is to act as a heuristic for people (a cognitive shortcut). On the one hand, if individuals are aware of the tendency for Republicans to use and be associated with the color red and the degree to which they know factual information about politics is not entirely relevant. On the other hand, those that pay no attention to politics whatsoever may be less likely to be aware of such associations. Additionally, those who express less interest in politics are also likely to identify as more moderate when asked about their ideology and partisan identity [@klar_independent_2016].

Additionally, I include a question about the respondent's sex assigned at birth, and importantly, about whether they have received a diagnosis of *any color blindness*. As some individuals may possess undiagnosed colorblindness, asking about their sex assists in covariate balance. I additionally include an open-ended question asking participants to describe their "first memory of a political event." The use of open-ended questions helps provide an attention check and identify duplicated responses for those spoofing IP addresses with a VPN [@kennedy_strategies_2021]. Given estimates using these exclusion criteria, I would expect that upwards of 40% of my original sample will fail these attention checks [@kennedy_strategies_2021], thus, the large sample size.

I then present participants with an instruction screen informing them of the task for the experiment. In the first trial of the experiment, I randomly presented participants with one of three possible yard signs. These yard signs are simple, with the text "Riley Ready to Lead" and a solid background color of either "Republican red," "Democratic Blue," or White.^[See [Appendix](#sec-2-a) to view all of the stimuli used in [Study 1](#study-1).] There is an added component to this, as well as in the other two trials.

Rather than use eye-tracking devices and software, I instead use Mouseview.js [see @anwyl-irvine_mouseviewjs_2022], which either blocks out or blurs a large portion of the participant's screen and encourages them to move their mouse to view different parts of the screen in isolation. As the participants move their cursor around the screen, it tracks the coordinates of the cursor along with the "dwell" time of the cursor in that particular coordinate. One primary benefit of Mouseview.js is that it allows researchers to field their experiments outside of a lab-based setting, while providing results that robustly correlate with the results from a design employing eye-tracking hardware [@anwyl-irvine_mouseviewjs_2022]. This allows researchers to rely less on student convenience samples, which are common with eye-tracking studies. For my design, I am particularly concerned about reliance on a student convenience sample due to variation in participants' ability to detect and process color in the U.S. population. This means that Mouseview.js is a handy tool for my study.

When viewing each yard sign in each of the three trials, there is a blur over a substantial portion of the screen. At any given point in time, participants can view only 8% of the image without an obstruction, which simulates the observation that we typically foveate on about 8% of our available visual field at any given time [@wedel_review_2008].^[I include a screenshot providing an example of what the participants are able to see with the blur included in the [Appendix](#sec-2-a).] Participants move their cursor to explore the yard sign. I allot 5000 ms to perform the exploration before the image goes away; this is to encourage a consistent and short duration to explore the image.^[In marketing research, some studies give participants about 6000 ms in eye-tracking studies to examine a brand and formulate an intention to purchase a product or not [@wedel_review_2008]. With Mouseview.js, a study examining the tool's correlation with optical responses to viewing disgust-and-pleasure evoking images uses 1000 ms; but is intended to be an extended amount of time [@anwyl-irvine_mouseviewjs_2022].]

After exploring the image in each trial, I ask participants whether they felt that the candidate supposedly owning the yard sign was a Democrat, a Republican, or Neither. After completing the three trials, I ask subjects to indicate their preference among the three images (one from each trial).

To ensure that participants have a standardized initial placement of their cursor, I display a blank page before viewing the yard sign that requires participants to click a "Next" button. Immediately after clicking "Next," participants are shown the yard sign. The goal is to ensure that variation in where participants explore the image is not dependent on a non-standard starting point for their cursor. I additionally utilize a gaussian blur for the overlay of the image, rather than a solid overlay obstructing the participant's view entirely. The gaussian blur allows participants to see a blurred visual field beyond the cursor. This allows participants to see enough to take purposeful action to explore blurred parts of the image that attract them [@anwyl-irvine_mouseviewjs_2022]. The use of the gaussian blur requires that participants use a web browser other than Safari because of a known issue [@anwyl-irvine_mouseviewjs_2022]. This will require participants to either switch browsers or to not participate in the study if they are using Safari at the time they are recruited by Prolific to participate in the study. As this requirement is enforced *before* joining the study, this should not have an effect on the number of excluded participants from my original sample.

There are two more trials that are much like the first trial. What is different between the two latter trials and the first is that I vary the amount of color that is on the yard sign. Examples of all of these yard signs are included in the [Appendix](#sec-2-a). For readers interested in the more technical details, the [Appendix](#sec-2-a) accompanying this chapter also includes a table of the data collected in the experiment -- @tbl-study-1-measures; reports full results, posterior predictive checks, and more detailed discussions for all of the models used in the remainder of this section.^[Unless noted otherwise, the reflection of uncertainty in the models are credible intervals at the 90% level. This should not be equated with the frequentist practice of reporting confidence intervals at the 95% level, as credible intervals at this level are often extremely unstable and, more importantly, are interpreted differently. For example, a credible interval at the 90% level can be interpreted as: the probability that the true effect is within that interval, given the data. This interval reflects the degree of uncertainty I have about my point estimate as opposed to being primarily used to interpret levels of statistical significance in a strict null hypothesis test. This contrasts with a confidence interval used in the frequentist framework which is interpreted as: upon repeated samples, 90% of the confidence intervals contain the true effect. To bridge methodological approaches, for those interpreting my credible intervals, I am using two-sided credible intervals here, at the 90% level, the probability that 0 would be the true value given the data would be less than 5% as it would be on one side of my posterior distribution.]

To address this first question, I use a couple of proxy measures for people's attention paid to the yard signs. I collect the first measure through recording the movement of the subjects' cursors along the x-and-y-axes. The second measure accounts for the amount of time participants' cursors linger over a particular coordinate on their screen. Presumably, the more participants explore the yard sign, the more they are finding visual information that they can use to evaluate the candidate. While these two measures are simple proxies and are not direct measures of "noticing the yard sign," they allow us to get a sense of the patterns by which people explore the yard signs across treatment conditions. This also bolsters my confidence in the degree to which variation between the treatments among different groups of participants is due to actual differences in how people processed these treatments, and not due to random chance.

Overall, it appears that people viewed the yard signs differently depending on which yard sign they were looking at. Now, I turn to exploring how features about the subjects themselves may shape the way that individuals interact and respond to this visual information on the yard signs. The data on the timing and cursor location is much more noisy than I had originally anticipated. However, as the tests of the next few questions demonstrate, the evidence supports the general conclusion of the systematic differences between individuals that saw either a red, white, or blue yard sign (despite holding everything else but the color constant).

To address the question of whether colors affect perceptions of the candidate and the yard sign, I ask participants to report whether they perceived the candidate to be a partisan -- either a Republican, Democrat, or as non-partisan -- immediately after viewing each image. Everything on the yard sign remains constant except for the color. As representations of ideology are associated with more than just political views, but also things like space [@mills_political_2016] and color [@losada_maestre_color_2022], differences between respondents on the perceived political affiliations of the candidate should be more than "by chance" differences.

@fig-h-2-predictions presents the predicted probabilities of an average subject reporting the candidate as either a Democrat, Independent, or Republican when viewing the different yard signs. The error bars reflect the high density credible interval for 90% of my posterior draws. Each bar reflects the predicted probability of an outcome -- listing the candidate as either a Democrat, Independent, or Republican. When interpreting the error bars, you will want to look at the difference in the error bars *between* yard sign treatments rather than *within* treatment. We see that the models indicate statistically meaningful differences between the yard sign treatments for each outcome. That is, the red and blue yard signs did shift people's perceptions of the partisanship of the candidate relative to the white yard sign. The credible intervals of these models are reported in the [Appendix](#sec-2-a) in @tbl-hypothesis-2-main which contains the rest of the results from these models.

Digging in a little bit more about what these perceptions were, the predicted probability that subjects perceive the candidate with a red yard sign to be Republican is around 70%, whereas it is only about 30% when viewing a white yard sign. When given a red yard sign, the predicted probability the candidate is marked to be an independent is about 30% and less than 5% to be a Democrat. However, when viewing a blue yard sign, the predicted probability of a subject perceiving the candidate to be a Democrat is at about 50% whereas it is less at about 10% with a white yard sign. When viewing a blue yard sign, the predicted probability of subjects perceiving the candidate to be a Republican is less than 20% and about 45% to be an Independent. The full results are reported in @tbl-hypothesis-2-main in the [Appendix](#sec-2-a).

As the consistency by which the parties have used either the color red or blue has increased since the 2000's, I suspect that the age of my subjects may moderate this relationship. That is, the tendency for subjects to label red yard signs as a Republican candidate versus a different color may be stronger for those who are younger as the consistency in branding is the political environment they have been around longer. I expect this particular relationship may be weaker for older subjects as they may place current branding strategies in a larger political context where this was not always the case.

I fit two ordered logistic regressions where the independent variables are indicator variables reporting whether the participant received the red (the first model) or they received the blue treatment (the second model). Those included in the baseline category for both models are those who received the white yard sign treatment. Since I believe that age may moderate this effect, I include an interaction term for both of these models that multiplies these treatment indicator variables with age.

I also explore whether these effects may be weaker for those with colorblindness. The results of this model is reported in [@tbl-colorblindness], which is located in the [Appendix](#sec-2-a). The results do not suggest that they are. However, given that I had a very small number of respondents reporting colorblindness in my sample, the lack of substantive or statistical effects may be due to the fact that I do not have enough participants with colorblindness in each group to compare.

I also test whether my effects here are sensitive to my choice to run two separate regression models -- one for each treatment. I fit a single ordered logistic regression with two indicator variables for the treatment coded as: 0 did not receive the red/blue treatment, 1 did receive treatment. The full results of this alternative model specification are included in [@tbl-robustness-alt-specification-h-2]. This means that the intercept term for the regression would be the predicted probability of the average respondent labeling the owner of the yard sign as either Democrat, Independent, or Republican. The conclusions I draw from this regression model do no0t change the ones I make here with @fig-h-2-predictions.

The design of the latter two trials in the study presents yard signs with mixtures of non-partisan and out-partisan colors. For example, I present primarily red (and presumably Republican) yard signs, but with some blue or white in them. I expect that the trials that use less "consistent" visual information demonstrate more uncertainty among respondents in their reported perceptions of the candidate's positions. Specifically, I expect that participants will perceive the yard signs that have both red and blue on them as more moderate, and that the higher proportion of the color red or blue among the two colors will lead respondents, on average, to be more likely to believe that the candidate leans more Republican or Democratic, respectively. This is not to say that I expect the results to change significantly, but rather, that the certainty by which I can predict individuals' guesses of the candidate's partisan affiliation will weaken.

The results from these models offer partial support for my expectations. First I see that yard signs using less blue are not as distinguishably Democratic as the white yard sign. In other words, voters are not meaningfully more likely to report that the yard sign with less blue on it belongs to a Democratic candidate versus a white and black yard sign. So the strength of using a blue yard sign to communicate alignment with the Democratic party weakens if you use less blue relative to white, or if you include red on it. With red yard signs, however, we see that voters still see them as belonging to Republican candidates, even when we reduce the amount of red on the yard sign and even when we include blue.

One potential reason for this may be the association of the American flag when using all three colors given the Republican party's tendency to condemn those who do not adhere to social norms about showing respect to the American flag. This, however, is untestable based on the design of the present study. However, what we do see is that there are bounds by which at least the color blue on a yard sign indicates to voters that the candidate is a Democrat. What remains unclear is whether these perceptions of ownership by a partisan candidate influence people's willingness to support that candidate or express favoritism toward them ($H_3$).

@fig-h-3-predictions presents the average predicted probability of supporting a candidate. The results fit with my expectations. First, I see that the predicted probability that a strong Democrat would vote for a candidate with a red yard sign is less than 10% (while for strong republicans it is at about 70%). Likewise, we see that strong Democrats much prefer a candidate with a blue yard sign than a candidate with a white yard sign, 80% relative to 40%, respectively. Strong Republicans are unlikely to vote for a candidate with a blue yard sign and a white yard sign, with their predicted probability of voting for either of those candidates below 30%.

The other analytical tasks do not examine expectations derived from the motivated reasoning portion of the model. That is, do people process in-group information faster than out-group information? There is, of course, evidence of these tendencies in political circumstances [@lodge_rationalizing_2013].

To examine whether motivated reasoning is also present in the processing of politically-relevant color, I examine the following: the difference in the amount of time between the start of viewing a stimulus and clicking "Next" to stop viewing the stimulus, among those who were viewing a presumed co-partisan yard sign relative to those viewing a presumed out-partisan yard sign. As the outcome is a count of the amount of time participants spent looking at a yard sign, I turn to a class of regression models that work well for explaining outcomes measured as count data.^[By the nature of my measure, I do not believe that participants can record a value of 0 as time elapsed is measured in milliseconds. Additionally, I expect over-dispersion, as I expect to observe a large bulk of participants recording a very small time elapsed, while for others I expect to record a quite large amount of elapsed time. In the appendix, @tbl-h-4-dispersion summarizes the time elapsed outcome variable; this demonstrates over-dispersion. Given this, rather than using a count model assuming a poisson-distributed outcome, I instead assume an outcome distributed as poisson-gamma.] I therefore fit two poisson-gamma models (sometimes referred to as negative binomial regressions). I am interested in testing whether time elapsed can be explained by not only the color of the yard sign that the subject is viewing, but also whether this effect is moderated by an individual's partisanship. I therefore include an interaction term between stimuli and the partisan identification of the respondent. Like before, as partisan identification is not randomly assigned, I control for confounds such as attention paid to politics, age, gender identity, and racial identity. @tbl-h-4 ([Appendix](#sec-2-a)) presents the full table of results for these two poisson-gamma regression models. @fig-h-4-ppc also indicates that the model does a quite good job at predicting the distribution of the time elapsed outcome variable.

[Study 1](#study-1) presents evidence in support of the idea that the colors red and blue matter to politics -- they convey information to people. The evidence from [Study 1](#study-1) also suggests that the information may shape the attitudes of voters. One thing that remains unclear is whether or not these patterns potentially have large enough impacts that they shape how campaigns market themselves. In informal conversations with a campaign marketing firm which specializes in campaign marketing, I was told:

@@

While this was good information to hear (that seemed to align with what I found in [Study 1](#study-1)), I wasn't sure how much this actually influenced strategy. It appears it does, however:

@@

Thus, the goal for the second piece of empirical evidence in this chapter is to explore whether there is any evidence -- outside of the "lab" -- suggesting that campaigns use the colors red and blue systematically. If the color of a yard sign does indeed matter, and its electoral effects depend on the perceptions of potential voters, we should expect that campaigns are less likely to place yard signs with a large amount of red or blue in more moderate districts. However, in districts where there is a strong partisan base, we should expect that the candidate of the popular party will use the party's color more, while the disadvantaged candidate will use their party's color less as a way to either play up or down their loyalty to their party.

To evaluate this, I begin by collecting data from the MIT Election Lab that reports the electoral results of the House of Representative elections from the years 1970-2020. The data here tell me which candidate got how many votes in each House election across the United States. With these data, I calculate the rolling 5-year average of the Democratic party's vote share for each district.

Next, I download around 1,000 yard signs from House elections from the years 2016 to 2024 from the website provided by the Center for American Politics and Design. I use the `OpenCV` library in `Python` to examine the proportion of the colors near the "Republican red"^[lower RGB values: 93, 9, 12; upper RGB values: 236, 69, 75] and "Democratic blue"^[lower RGB values: 0, 18, 26; upper RGB values [102, 212, 255]] in each of these yard signs.

@@

With the data on the proportion of each of the yard signs using the red and blue colors, I merge this data with the MIT Election Lab Data. This leaves me with data from the House elections between 2018 and 2020. While the data are limited in the number of elections they reflect, they allow me to hold the district composition constant, as redistricting occurs after the Census every 10 years. However, each district and state is different in their laws surrounding elections, and are different in terms of political context. I want to examine whether there are meaningful differences between districts that are partisan "strongholds" (Democrats or Republicans tend to get elected in a district) relative to districts that are much less so. To do this, I include a model with intercepts with no pooling. That is, a model that includes an intercept for each state-district unit. Though my outcome of interest (proportion of the yard sign that is red or blue) are continuous (it can take any real number, so not an integer), the outcome is bounded between 0 and 1. While the traditional approach may be to use ordinary least squares, such a technique often can lead to poor model fit, introduce systematic error, and lead to results that are quite unstable. To not bog down readers with a lot of technical jargon, the model that I end up using is called an ordered beta regression which allows me to efficiently and accurately predict outcomes that can be 0 or 1, or any real number between 0 and 1 [@kubinec_ordered_2022]. For a more technical discussion about why I choose this over other techniques in the beta regression family, please see the [Appendix](#sec-2-a).

Once I have fit these models, I use the `marginaleffects` [@arel-bundock_marginaleffects_2022] package to calculate the grand mean of the predicted proportion of the use of the color red or blue for these yard signs, as explained by the 5-year rolling average Democratic candidates' vote share. The grand mean here reflects that I am not showing the predicted proportion of the use of these outcomes for each state-district, but rather that I am averaging across all districts to provide a single line in my figure. These results are included in @fig-study-2-pred.

@@

The results demonstrate that districts where Democrats are historically more successful see yard signs with a lower proportion of red on them, while those same districts see more blue on them. Again, these results do not claim that campaigns are doing this either because voters are encouraging them to take this strategy, or that they are doing it as a way to proactively get voters to do it. To tease out which of those questions is the right one requires more sophisticated approaches and data that could allow me to leverage redistricting (to see how these outcomes change due to some unrelated, external shock). The data available to me do not allow me to do that at this time. Rather, these data provide evidence consistent with the story I have relayed, and motivate further the study of color as it relates to politics.

>>>>

---
title: |
  Does Color Convey Political Information?
short-title: Color as Political Information
published: "Manuscript draft: please do not share without author permission."
code-repo: "For replication, go to: <https:://github.com/DamonCharlesRoberts/book_project>."
author:
  - name: Damon C. Roberts
    email: damon.roberts-1@colorado.edu
    orcid: 0000-0002-4360-3675
    title: PhD Candidate
    affiliations:
      - id: CU
        name: University of Colorado Boulder
        department: Political Science
        address: 333 UCB
        city: Boulder
        region: CO 
        postal-code: 80309-0333
abstract: |
  Do the colors red and blue convey political information? If so, how do they shape candidate preferences and vote choice? Political psychologists rely on interdisciplinary approaches to construct theories of how we process political information. Likewise, those who study visual political information come from a number of disciplinary traditions to study how politics may be conveyed visually. As these groups of scholars are working in parallel, they have yet to fully integrate their work to construct a theory explaining the cognitive and behavioral consequences of visual information on political attitude expression and behavior. In this chapter, I present a snap-judgment model of visual information processing to demonstrate how visual information efficiently conveys substantive political information. Specifically, I focus on a simple but obvious case that we only have anecdotal and descriptive evidence of: the association between Republicans with the color red and Democrats with the color blue. First, I use supervised machine learning that does a pixel-by-pixel classification of the color on yard signs used by campaigns for the U.S. House of Representatives from 2016 to 2020 and merge this with election return data from the MIT Election Labs. From these data, I find that systematic descriptive evidence of campaigns using more red in their yard signs in districts that are safe for Republicans and campaigns using more blue in their yard signs in districts that are safe for Democrats. Next, I turn to an experiment to examine whether these associations of the colors red and blue with Republicans and Democrats exist among voters and whether it shapes attitude expression and behavior. I find that these associations of the color red with Republicans and the color blue with Democrats do exist among the public and that these associations are strong enough to shape support for political candidates. This chapter advances our understanding in a few areas: that visual information is a pre-conscious form of information that we should integrate with our models of information processing that consider more traditional forms of political information, provides a cognitive mechanism to explain how the public have come to associate the colors red and blue with the two dominant political parties, as well as demonstrates that these cognitive processes are capable of influencing behavior as important as vote choice.
additional-info: |
  I have no known conflict of interest to disclose.
keywords:
  - visual information
  - information processing
  - partisanship
date: today
bibliography: "../../assets/candidate_support_chapter_references.bib"
format:
  hikmah-pdf:
    cite-method: citeproc
    latex-output-dir: ../out/
    geometry:
      - top=1in
      - bottom=1in
      - left=1in
      - right=1in
      - heightrounded
    linestretch: 2
    fontsize: 12pt
    appendix-style: default
    include-in-header:
      text: |
        \usepackage{pdflscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
linestretch: 2
execute:
  echo: false
  warning: false
  message: false
params:
  replicate: false
---

```{r}
#| label: r-setup-block

@@

In [Study 1](#study-1), I find descriptive evidence suggesting that campaigns use the colors red and blue on their yard signs at different rates depending on the success of Republicans and Democrats according to previous election returns. It is unclear, however, whether these patterns are based on elite perceptions of how to communicate partisanship to voters or whether campaigns have successfully identified a cognitive process linking the colors red and blue with Republicans and Democrats. Accordingly, in [Study 2](#study-2), I conduct an online experiment that examines whether the public notice color and whether their attitudes are shaped by the presence of "Republican red" and "Democratic blue" on the yard signs of a political candidate. I find that these associations do exist among the public. Specifically, individuals consistently and strongly associate Republicans with the color red and the Democrats with the color Blue. As a result, partisan individuals' willingness to vote for a candidate can be dramatically different depending on whether the color used in the yard sign aligns with their partisan identification.

While these two studies highlight endogenous processes between campaigns and voters, to be clear, I am not arguing that either of these patterns only one way. In fact, I believe that there is a necessary feedback loop between the parties and their bases of support. Voters support candidates who signal strong allegiance to the party to increase their chances for electoral success [@utych_voter-centric_2020], and the parties are motivated to distinguish themselves [@lee_insecure_2016;@clifford_how_2020] as a result of their high levels of ideological and affective polarization [@dietrich_using_2021;@enders_issues_2021]. I leave teasing out directionality for subsequent analyses.

@@

In [Study 1](#study-1) I start with testing $H_5$. In some informal conversations that I had with political campaign professionals, it became clear that the descriptive evidence connecting the colors red and blue with Republicans and Democrats is still very much based on experience.

@@

It appears that campaigns recognize that these are not simply patterns that journalists [see @elving_color_2014] and researchers [see @williams_whats_2022] have picked up on, bit that they are associations held by voters and campaigns respond accordingly:

@@

Though it appears that at least one consulting firm of campaign marketing professionals recognize the potential value of using the colors red and blue to cue to the partisanship of a candidate, there is still significant uncertainty about how widespread these patterns are with political campaigns in the United States and whether voters actually do make these associations. The first piece of empirical evidence in this chapter examines whether we observe systematic patterns between campaigns in the use of the colors red and blue on their yard signs. [Study 2](#study-2) then takes my predictions from the snap-judgment model to examine whether these associations exist among voters and what the consequences are.

To gather evidence generalizing the belief that campaigns strategically use the colors red and blue to tap into possible associations of these colors with partisanship, I began by collecting data from the MIT Election Lab that reports the electoral results of the House of Representative elections from the years 1970-2020. The data tell me which candidate got how many votes in each House election across the United States. With these data, I calculate the rolling 5-year average of the Democratic party's vote share for each district.

Next, I downloaded around 1,000 yard signs from House elections from the years 2016 to 2024 from the website provided by the Center for American Politics and Design. I used the `OpenCV` library in `Python` to examine the proportion of the colors near the "Republican red"^[lower RGB values: 93, 9, 12; upper RGB values: 236, 69, 75] and "Democratic blue"^[lower RGB values: 0, 18, 26; upper RGB values [102, 212, 255]] in each of these yard signs.

@@

With the data on the proportion of each of the yard signs using the red and blue colors, I merged these data with the MIT Election Lab Data. This left me with data from the House elections between 2018 and 2020. While the data are limited in the number of elections they reflect, they allow me to hold the district composition constant, as redistricting occurs after the Census every 10 years. However, each district and state is different in their laws surrounding elections, and are different in terms of political context. I wanted to examine whether there are meaningful differences between districts that are partisan "strongholds" (Democrats or Republicans tend to get elected in a district) relative to districts that are much less so. To do this, I included a model with intercepts with no pooling. That is, a model that includes an intercept for each state-district unit. Though my outcome of interest (proportion of the yard sign that is red or blue) are continuous (it can take any real number, so not an integer), the outcome is bounded between 0 and 1. While the traditional approach may be to use ordinary least squares, such a technique often can lead to poor model fit, introduce systematic error, and lead to results that are quite unstable. To not bog down readers with a lot of technical jargon, the model that I end up using is called an ordered beta regression which allows me to efficiently and accurately predict outcomes that can be 0 or 1, or any real number between 0 and 1 [@kubinec_ordered_2022]. For a more technical discussion about why I choose this over other techniques in the beta regression family, please see the [Appendix](#sec-2-a).

Once I fit these models, I used the `marginaleffects` [@arel-bundock_marginaleffects_2022] package to calculate the grand mean of the predicted proportion of the use of the color red or blue for these yard signs, as explained by the 5-year rolling average Democratic candidates' vote share. The grand mean here reflects that I am not showing the predicted proportion of the use of these outcomes for each state-district, but rather that I am averaging across all districts to provide a single line in my figure. These results are included in @fig-study-2-pred.

@@

The results demonstrate that districts where Democrats are historically more successful see yard signs with a lower proportion of red on them, while those same districts see more blue on them. Again, these results do not claim that campaigns are doing this either because voters are encouraging them to take this strategy, or that they are doing it as a way to proactively get voters to make these associations. To tease out which of those questions is the right one requires more sophisticated approaches and data that could allow me to leverage redistricting (to see how these outcomes change due to some unrelated, external shock). The data available to me do not allow me to do that at this time. Rather, these data provide evidence consistent with the story I have relayed and the one provided to me in my informal conversations with practitioners.

The purpose of [Study 2](#sec-study-2) is to examine whether the color choices on an almost ubiquitous form of campaign branding -- yard signs -- influence people's perceptions of political candidates. Evidence suggests that yard signs *do* matter in shaping political attitudes [@makse_politics_2019], vote intentions [@makse_politics_2019], and even electoral outcomes [@green_effects_2016]. As they are simple, cheap, intrusive, and common forms of campaign branding, they provide a conservative test of the effects of color in campaign branding.

Using yard signs in this study, I test the first four hypotheses I derived from the snap-judgment model. First, I tested the claim that individuals do indeed notice the color of electoral yard signs. Next, I tested the claim that these colors that individuals detect influence their evaluations of the yard sign and the candidate represented on it. I then tested the claim that the effects on perception are moderated by how consistent the color is with the more complex information displayed on the yard sign. Finally, [Study 2](#study-2) closes with an examination of whether positive information that contains consistency is processed more quickly than negative or inconsistent information. That is, do partisans quickly detect and encode information from a clearly co-partisan political candidate?

I recruited participants from Prolific.[^color_influence_pre-reg_anon-3] After providing informed consent to participate in the study, Prolific redirects subjects to Pavlovia,[^color_influence_pre-reg_anon-4] where individuals are provided with a demographics and political attitudes questionnaire. As I am concerned about priming effects introduced by these questions, while I am simultaneously concerned about bias introduced by post-treatment control [@montgomery_how_2018], I randomly selected half of the participants to receive the questionnaire post-treatment and the other half receive it pre-treatment. This questionnaire includes common questions about the participant's ascriptive characteristics, and about the participant's political ideology, partisan identification, interest in politics, patriotism, and political knowledge.

I included those questions due to expectations that they may act as confounds for my hypotheses. I included some questions collecting information on participants' ascriptive and descriptive characteristics such as age, education, gender identity, and racial identity, as a number of these correlates with partisan identification [see @campbell_american_1969;@mason_uncivil_2018]. Though political knowledge does predict the strength of an individual's partisan identification [@lodge_rationalizing_2013], attention paid to politics is likely a more important confounding variable for my application. The purpose of color in politics is to act as a heuristic for people (a cognitive shortcut). On the one hand, if individuals are aware of the tendency for Republicans to use and be associated with the color red and the degree to which they know factual information about politics is not entirely relevant. On the other hand, those that pay no attention to politics whatsoever may be less likely to be aware of such associations. Additionally, those who express less interest in politics are also likely to identify as more moderate when asked about their ideology and partisan identity [@klar_independent_2016].

Additionally, I included a question about the respondent's sex assigned at birth, and importantly, about whether they have received a diagnosis of *any color blindness*. As some individuals may possess undiagnosed colorblindness, asking about their sex assists in covariate balance. I additionally included an open-ended question asking participants to describe their "first memory of a political event." The use of open-ended questions helps provide an attention check and identify duplicated responses for those spoofing IP addresses with a VPN [@kennedy_strategies_2021]. Given estimates using these exclusion criteria, I expected that upwards of 40% of my original sample will fail these attention checks [@kennedy_strategies_2021], thus, the large target sample size.

I then presented participants with an instruction screen informing them of the task for the experiment. In the first trial of the experiment, I randomly presented participants with one of three possible yard signs. These yard signs are simple, with the text "Riley Ready to Lead" and a solid background color of either "Republican red," "Democratic Blue," or White.^[See [Appendix](#sec-2-a) to view all of the stimuli used in [Study 1](#study-1).] There is an added component to this, as well as in the other two trials.

Rather than use eye-tracking devices and software, I instead used Mouseview.js [see @anwyl-irvine_mouseviewjs_2022], which either blocks out or blurs a large portion of the participant's screen and encourages them to move their mouse to view different parts of the screen in isolation. As the participants move their cursor around the screen, it tracks the coordinates of the cursor along with the "dwell" time of the cursor in that particular coordinate. One primary benefit of Mouseview.js is that it allows researchers to field their experiments outside of a lab-based setting, while providing results that robustly correlate with the results from a design employing eye-tracking hardware [@anwyl-irvine_mouseviewjs_2022]. This allows researchers to rely less on student convenience samples, which are common with eye-tracking studies. For my design, I am particularly concerned about reliance on a student convenience sample due to variation in participants' ability to detect and process color in the U.S. population. This means that Mouseview.js was a handy tool for my study.

When viewing each yard sign in each of the three trials, there is a blur over a substantial portion of the screen. At any given point in time, participants can view only 8% of the image without an obstruction, which simulates the observation that we typically foveate on about 8% of our available visual field at any given time [@wedel_review_2008].^[I include a screenshot providing an example of what the participants were able to see with the blur included in the [Appendix](#sec-2-a).] Participants move their cursor to explore the yard sign. I alloted 5000 ms to perform the exploration before the image disappears; I did this to encourage a consistent and short duration to explore the image.^[In marketing research, some studies give participants about 6000 ms in eye-tracking studies to examine a brand and formulate an intention to purchase a product or not [@wedel_review_2008]. With Mouseview.js, a study examining the tool's correlation with optical responses to viewing disgust-and-pleasure evoking images uses 1000 ms; but is intended to be an extended amount of time [@anwyl-irvine_mouseviewjs_2022].]

After exploring the image in each trial, I asked participants whether they felt that the candidate supposedly owning the yard sign was a Democrat, a Republican, or Neither. After completing the three trials, I asked subjects to indicate their preference among the three images (one from each trial).

To ensure that participants have a standardized initial placement of their cursor, I displayed a blank page before viewing the yard sign that requires participants to click a "Next" button. Immediately after clicking "Next," participants were shown the yard sign. The goal is to ensure that variation in where participants explore the image is not dependent on a non-standard starting point for their cursor. I additionally utilized a gaussian blur for the overlay of the image, rather than a solid overlay obstructing the participant's view entirely. The gaussian blur allowed participants to see a blurred visual field beyond the cursor. This allowed participants to see enough to take purposeful action to explore blurred parts of the image that attract them [@anwyl-irvine_mouseviewjs_2022]. The use of the gaussian blur required that participants use a web browser other than Safari because of a known issue [@anwyl-irvine_mouseviewjs_2022]. This required participants to either switch browsers or to not participate in the study if they are using Safari at the time they are recruited by Prolific to participate in the study. As this requirement is enforced *before* joining the study, this should not have an effect on the number of excluded participants from my original sample.
